Spá: a web-based viewer for text mining in Evidence Based Medicine

### 2014-04-24 documentation/paper.tex
Remove redundant packages
Minor fixes to paper

% Created 2014-04-14 Mon 18:10
\documentclass[runningheads,a4paper]{llncs}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{wrapfig}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage[multiple]{footmisc}
\usepackage{microtype}
\usepackage{acronym}
\tolerance=1000
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{listings}

@@

\acrodef{ebm}[EBM]{Evidence-based Medicine}
% I don't think thi sneeds to be capitalized -- bcw
\acrodef{ml}[ML]{machine learning}
\acrodef{pdf}[PDF]{Portable Document Format}
\acrodef{cdsr}[CDSR]{Cochrane Database of Systematic Reviews}
\acrodef{svm}[SVM]{Support Vector Machine}

@@

Imposing structure on full-text documents (e.g., identifying specific sentences of interest) is an important and practical task in natural language processing and machine learning, because doing so manually is laborious and costly.
\emph{Systematic reviews} are an instructive example.
Such reviews aim to answer clinical questions by exhaustively identifying and extracting information from all published relevant studies.
They are fundamental tools in \ac{ebm} \cite{sackett1996}.
The end product of a systematic review is an exhaustive synthesis (textual and statistical) of the current evidence.
To produce these, reviewers must manually extract data on trial design and conduct, intervention effectiveness, adverse effects, and potential biases from the free-text of published studies.
These extraction tasks can be semi-automated using machine learning approaches.
As an example, here we consider the task of bias assessment, in which a reviewers assess, e.g., whether study participants and personnel were properly blinded.

Assessing this is a time-consuming task: a single trial can take a domain expert around ten minutes \cite{Hartling2011}.
Making matters worse, a single review may typically include several dozen trials, and due to low rates of reviewer agreement it is regarded as best practice to have each study assessed twice by independent reviewers who later come to a consensus \cite{Hartling2009}.
Machine learning methods provide the machinery to automate such extractions; as they can effectively impose the desired structure onto \acp{pdf}. 
But if such technologies are to be practically useful, we need tools that visualize these model predictions and annotations.
Here we describe Spá, which aspires to realize this aim.

Spá is an open-source, web-based tool that can incorporate state-of-the-art machine learning predictors to automatically annotate \ac{pdf} articles.
As a practical demonstration of this technology, we have built a machine learning system that automatically annotates \acp{pdf} to aid \ac{ebm}.
% maybe?
Specifically, this system leverages a multi-task model that simulteanously assesses the risk of biases across several domains.
This tool is unique in that it leverages state-of-the-art \ac{ml} models applied to full-text articles to assist practitioners of \ac{ebm}.

@@

While our application of interest is \ac{ebm}, we emphasize that the visualization tool can be used for any domain in which one wants to annotate \acp{pdf}.
Thus the contribution of this work is two-fold, as we present:
(1) a practical tool that incorporates machine learning to help researchers rapidly assess the risk of biases in published biomedical articles, and,
(2) a general open-source web tool for visualizing the predictions of trained models from full-text articles.
These contributions are described further in Sections \ref{section:EBM-ML} and \ref{section:architecture}.

@@

\section{Future work}
We have presented a web-based tool for visualization of annotations and summarizations for \ac{pdf} documents.
It allows users to see the results from machine learning predictions for a specific document.
We have demonstrated the use of this system within the context of \ac{ebm} by automatically extracting potential risks of bias (and supporting sentences) via state-of-the-art ML methods.

More generally, we believe the tool to be potentially useful for a much wider range of text mining and machine learning applications.
To increase the generality of the tool we are developing a pluggable system for processing topologies, allowing developers to quickly plug in new ML systems for automated \ac{pdf} annotation.
Furthermore, we are working to allow users to save selected annotations, possibly embedded within the document itself, for sharing and off-line use.
The vision is to have an extensible system for semi-automated (machine assisted) screening, data extraction and data summarization for EBM, and to allow rapid development of similar systems for other domains.

>>>>

% Created 2014-04-14 Mon 18:10
\documentclass[runningheads,a4paper]{llncs}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage[multiple]{footmisc}
\usepackage{acronym}
\tolerance=1000
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{listings}

@@

\acrodef{ebm}[EBM]{Evidence-based Medicine}
\acrodef{ml}[ML]{machine learning}
\acrodef{pdf}[PDF]{Portable Document Format}
\acrodef{cdsr}[CDSR]{Cochrane Database of Systematic Reviews}
\acrodef{svm}[SVM]{Support Vector Machine}

@@

Imposing structure on full-text documents (e.g., identifying specific sentences of interest) is an important and practical task in natural language processing and machine learning, because doing so manually is laborious and costly.
\emph{Systematic reviews} are an instructive example.
Such reviews aim to answer clinical questions by exhaustively identifying and extracting information from all published relevant studies.
They are fundamental tools in \ac{ebm} \cite{sackett1996}.
The end product of a systematic review is an exhaustive synthesis (textual and statistical) of the current evidence.
To produce these, reviewers must manually extract data on trial design and conduct, intervention effectiveness, adverse effects, and potential biases from the free-text of published studies.
These extraction tasks can be semi-automated using machine learning approaches.
As an example, here we consider the task of bias assessment, in which reviewers assess, e.g., whether study participants and personnel were properly blinded.

Assessing this is a time-consuming task: a single trial can take a domain expert around ten minutes \cite{Hartling2011}.
Making matters worse, a single review may typically include several dozen trials, and due to low rates of reviewer agreement it is regarded as best practice to have each study assessed twice by independent reviewers who later come to a consensus \cite{Hartling2009}.
\ac{ml} methods provide the machinery to automate such extractions; as they can effectively impose the desired structure onto \acp{pdf}.
But if such technologies are to be practically useful, we need tools that visualize these model predictions and annotations.
Here we describe Spá, which aspires to realize this aim.

Spá is an open-source, web-based tool that can incorporate state-of-the-art \acl{ml} predictors to automatically annotate \ac{pdf} articles.
As a practical demonstration of this technology, we have built a machine learning system that automatically annotates \acp{pdf} to aid \ac{ebm}.
% maybe?
Specifically, this system leverages a multi-task model that simulteanously assesses the risk of biases across several domains.
This tool is unique in that it leverages state-of-the-art \acl{ml} models applied to full-text articles to assist practitioners of \ac{ebm}.

@@

While our application of interest is \ac{ebm}, we emphasize that the visualization tool can be used for any domain in which one wants to annotate \acp{pdf}.
Thus the contribution of this work is two-fold, as we present:
(1) a practical tool that incorporates machine learning to help researchers rapidly assess the risk of biases in published biomedical articles, and,
(2) a general open-source web tool for visualizing the predictions of trained models from full-text articles.
These contributions are described further in sections \ref{section:EBM-ML} and \ref{section:architecture}.

@@

\section{Future work}
We have presented a web-based tool for visualization of annotations and summarizations for \ac{pdf} documents.
It allows users to see the results from machine learning predictions for a specific document.
We have demonstrated the use of this system within the context of \ac{ebm} by automatically extracting potential risks of bias (and supporting sentences) via state-of-the-art \ac{ml} methods.

More generally, we believe the tool to be potentially useful for a much wider range of text mining and machine learning applications.
To increase the generality of the tool we are developing a pluggable system for processing topologies, allowing developers to quickly plug in new \ac{ml} systems for automated \ac{pdf} annotation.
Furthermore, we are working to allow users to save selected annotations, possibly embedded within the document itself, for sharing and off-line use.
The vision is to have an extensible system for semi-automated (machine assisted) screening, data extraction and data summarization for EBM, and to allow rapid development of similar systems for other domains.
